Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### 2023-12-06 doc/manuscript/translocation.qmd
Quarto auto-formats

To determine whether MYL frog populations show an evolutionary response to Bd, we compared frog genomes between populations with contrasting histories of Bd exposure.
Specifically, we compared frog genomes sampled in 4 populations that have not yet experienced a Bd-caused epizootic ("naive") [@zhou2015] versus in 5 populations that experienced a Bd epizootic during the past several decades and have since recovered to varying degrees ("recovering"; @fig-allelefrequencies) [@knapp2016; @vredenburg2010].
Bd-exposure histories of the nine study populations are based on 10-20 years of visual encounter surveys and Bd surveillance using skin swabbing [e.g., @knapp2016; @zhou2015; @wilber2022].
Naive populations are characterized by large numbers of adults (i.e, typically 1000s), Bd prevalence that is generally 0% except during occasional Bd failed invasions [during which Bd loads remain very low; wilber2022], and no history of Bd epizootics since we first first surveyed these populations in the late 1990s and early 2000s [@zhou2015]. 
In contrast, recovering populations exist in an enzootic state [@briggs2010], characterized by relatively small numbers of adults (generally < 100), high Bd prevalence (often > 80%; [@knapp2011], and, in adults, moderate Bd loads that are typically well below the level expected to cause mortality [@vredenburg2010]. 
Naive and recovering populations can be identified unambiguously using these differences in Bd prevalence and load. 
(See SI Appendix for additional details regarding the study design.)  
<!--In response to Editor - Comment 3, mention that there is no dispersal between extant populations due to presence of fish and fish-induced habitat fragmentation?-->

@@

| **Parameter**                                                             | **Value**            | **Source**                                                                        |
|:--------------------------|:------------------|:-------------------------|
| $\sigma_{L_1}$, Yearly survival probability of year 1 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{L_2}$, Yearly survival probability of year 2 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{L_3}$, Yearly survival probability of year 3 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{J_1}$, Yearly survival probability of year 1 juveniles           | Varies yearly        | Varies. Bounds estimated from field data, observations, natural history knowledge |
| $\sigma_{J_2}$, Yearly survival probability of year 2 juveniles           | 0.5                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{A_R}$, Yearly survival probability of naturally recruited adults | Varies by population | Estimated from mark-recapture studies                                             |
| $\sigma_{A_T}$, Yearly survival probability of translocated adults        | Varies by population | Estimated from mark-recapture studies                                             |
| $p_{L_1}$, Probability of a year 1 tadpoles remaining as a tadpoles       | 1                    | Estimated from field data, observations, natural history knowledge                |
| $p_{L_2}$, Probability of a year 2 tadpoles remaining as a tadpoles       | 0.25                 | Estimated from field data, observations, natural history knowledge                |
| $p_{J_1}$, Probability of a year 1 juvenile remaining as a juvenile       | 0.25                 | Estimated from field data, observations, natural history knowledge                |
| $p_F$, Probability of a adult female reproducing in a year                | 0.5                  | Could be as high at 1, based on field observations                                |
| $F$, Number of surviving eggs produced by an adult female                 | 100                  | From observations of captive frogs                                                |

>>>>

To determine whether MYL frog populations show an evolutionary response to Bd, we compared frog genomes between populations with contrasting histories of Bd exposure.
Specifically, we compared frog genomes sampled in 4 populations that have not yet experienced a Bd-caused epizootic ("naive") [@zhou2015] versus in 5 populations that experienced a Bd epizootic during the past several decades and have since recovered to varying degrees ("recovering"; @fig-allelefrequencies) [@knapp2016; @vredenburg2010].
Bd-exposure histories of the nine study populations are based on 10-20 years of visual encounter surveys and Bd surveillance using skin swabbing [e.g., @knapp2016; @zhou2015; @wilber2022].
Naive populations are characterized by large numbers of adults (i.e, typically 1000s), Bd prevalence that is generally 0% except during occasional Bd failed invasions \[during which Bd loads remain very low; wilber2022\], and no history of Bd epizootics since we first first surveyed these populations in the late 1990s and early 2000s [@zhou2015].
In contrast, recovering populations exist in an enzootic state [@briggs2010], characterized by relatively small numbers of adults (generally \< 100), high Bd prevalence (often \> 80%; [@knapp2011], and, in adults, moderate Bd loads that are typically well below the level expected to cause mortality [@vredenburg2010]. Naive and recovering populations can be identified unambiguously using these differences in Bd prevalence and load. (See SI Appendix for additional details regarding the study design.)\
<!--In response to Editor - Comment 3, mention that there is no dispersal between extant populations due to presence of fish and fish-induced habitat fragmentation?-->

@@

| **Parameter**                                                             | **Value**            | **Source**                                                                        |
|:--------------------------------------------------------------------------|:---------------------|:----------------------------------------------------------------------------------|
| $\sigma_{L_1}$, Yearly survival probability of year 1 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{L_2}$, Yearly survival probability of year 2 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{L_3}$, Yearly survival probability of year 3 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{J_1}$, Yearly survival probability of year 1 juveniles           | Varies yearly        | Varies. Bounds estimated from field data, observations, natural history knowledge |
| $\sigma_{J_2}$, Yearly survival probability of year 2 juveniles           | 0.5                  | Estimated from field data, observations, natural history knowledge                |
| $\sigma_{A_R}$, Yearly survival probability of naturally recruited adults | Varies by population | Estimated from mark-recapture studies                                             |
| $\sigma_{A_T}$, Yearly survival probability of translocated adults        | Varies by population | Estimated from mark-recapture studies                                             |
| $p_{L_1}$, Probability of a year 1 tadpoles remaining as a tadpoles       | 1                    | Estimated from field data, observations, natural history knowledge                |
| $p_{L_2}$, Probability of a year 2 tadpoles remaining as a tadpoles       | 0.25                 | Estimated from field data, observations, natural history knowledge                |
| $p_{J_1}$, Probability of a year 1 juvenile remaining as a juvenile       | 0.25                 | Estimated from field data, observations, natural history knowledge                |
| $p_F$, Probability of a adult female reproducing in a year                | 0.5                  | Could be as high at 1, based on field observations                                |
| $F$, Number of surviving eggs produced by an adult female                 | 100                  | From observations of captive frogs                                                |
